---
title: "README"
output: html_document
---

```{r echo = F, message = F}
library(dplyr)
library(tidyr)
library(ggplot2)
source('R/funcs.R')
```


#### Files

All data created in `R\dat_proc.R`.  Source data in the ignore folder were created elsewhere (carpveg proj).

* `fish_all.RData` DNR trapnet/gillnet fish data, one unique record per fish.  Only Carp and bullhead species are identified, all other species labelled as 'other'. Data are only for July, August, September.  Includes true zeroes. 

* `fish_dat.RData` Same as `fish_all.RData` but data are converted to CPUE.  CPUE estimated as number of fish divided by effort, unique to species, date, lake etc.  CPUE was esimated separately for trapnet, gillnet, then both were summed. 

* `fishveg_dat.RData` combined fisheries and veg data, veg data summarized by total rich and subm rich for each lake.  Fish and veg data combined to be +/- one year of each other.  

* `veg_dat.RData` DNR veg transect data from 1992 to present. Format is dow, date, transect, species, and abundance category.  NULL abundance entries are not removed, these are species in the survey but not observed on a transect.  Note that there were no lakes in the dataset that had zero veg.  

#### EDA

```{r}

data(fishveg_dat)

# cpue histograms

toplo <- select(fishveg_dat, matches('^BHD|^CAP')) %>% 
  gather %>% 
  filter(value > 0)
 
ggplot(toplo, aes(x = key)) + 
  geom_bar() + 
  facet_wrap(~key) + 
  theme_bw() +
  scale_x_continuous(limits = c(-10, 200))

```

